---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, set-up, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  out.width = "100%"
)
remotes::install_local()
```

# Australian weather data

This package produces graphs of temperatures in summer and winter for Australian
locations using data downloaded from the
[Australian Bureau of Meteorology](http://www.bom.gov.au/?ref=logo) (BoM).

Daily maximum and minimum temperature measurements can be downloaded for weather
stations in Australia from the
[Climate Data Online section](http://www.bom.gov.au/climate/data/) of the BoM
website.

The package includes pre-downloaded datasets for the main airport location in
the capital city of each Australian state and territory:

* Adelaide

* Brisbane

* Canberra

* Darwin

* Hobart

* Melbourne

* Perth

* Sydney

These can be loaded via `data('Adelaide')`, for example. Otherwise, use
`download_temperatures()` to download the most recent temperatures for these
cities or for other locations.

## Installation

You can install the development version of `austemp` from
[GitHub](https://github.com/a-s-russo/austemp.git) with:

``` r
# install.packages("devtools")
devtools::install_github("a-s-russo/austemp")
```

## Example

Download summer and winter data for Adelaide:

(Approximate download date & time is
`r format(Sys.time(), "%d %b %Y @ %H%p %Z")`.)

```{r example-download, message=FALSE}
# Load the package
library(austemp)

# Download temperature data
station <- '023034' # Station number for Adelaide Airport
product_max <- '122' # Product number for maximum temperatures
product_min <- '123' # Product number for minimum temperatures
URLpart1 <-
  'http://www.bom.gov.au/jsp/ncc/cdio/weatherData/av?p_nccObsCode='
URLpart2 <-
  '&p_display_type=dailyDataFile&p_startYear=&p_c=&p_stn_num='
max_temp_URL <- paste0(URLpart1, product_max, URLpart2, station)
min_temp_URL <- paste0(URLpart1, product_min, URLpart2, station)
Adelaide <-
  download_temperatures(URLs = c(max_temp_URL, min_temp_URL))
```

(Alternatively, use the pre-downloaded dataset for Adelaide via
`data('Adelaide')` as mentioned above.)

The downloaded data can then be graphed:

```{r example-plot-summer, message=FALSE, out.height="100%", out.width="100%"}
# Load the package
library(austemp)
# Graph summer temperatures
plot_temperatures(
  data = Adelaide,
  season = 'summer',
  location = 'Adelaide Airport',
  thresholds = c(30, 35, 40),
  start_year = 2001,
  end_year = 2023
)
```

(There are missing temperatures in 2024 given the download date above. There are
also no temperatures for 29 February in non-leap years obviously.)

```{r example-plot-winter, message=FALSE, out.height="100%", out.width="100%"}
# Graph winter temperatures
plot_temperatures(
  data = Adelaide,
  season = 'winter',
  location = 'Adelaide Airport',
  thresholds = c(0, 3, 5),
  start_year = 2000,
  end_year = 2023
)
```

(There are no temperatures for 2024 yet given the download date above.)